
\documentclass[11pt,twoside,notitlepage]{report}

\usepackage{indentfirst}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{url}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage{Sweave}

\usepackage{ifthen}
\newboolean{techrept}
\setboolean{techrept}{true}
\newboolean{submit}
\setboolean{submit}{false}

\input{macros}

% remove room for headers, add to textheight

  \addtolength{\textheight}{\headheight}
  \addtolength{\textheight}{\headsep}
  \setlength{\headheight}{0 pt}
  \setlength{\headsep}{0 pt}

% adjust so right and left hand pages have different margins
% not sure why these particular numbers were picked or if they
% still make sense

  % \showthe\evensidemargin
  % \showthe\oddsidemargin
  % \showthe\textwidth

  % \evensidemargin  15 pt
  % \oddsidemargin 61.5 pt
  % \textwidth 392.5 pt

  \evensidemargin 28.75 pt
  \oddsidemargin 75.25 pt
  \textwidth 365 pt

\begin{document}

  \vspace*{0.9375in}
  \begin{center}
    {\bfseries Aster Models for Life History Analysis} \\
    By \\
    Charles J. Geyer, Stuart Wagenius, and Ruth G. Shaw \\
    Technical Report No.~644 \\
    School of Statistics \\
    University of Minnesota \\
%       April 20, 2005 \\
%       \today
    September 5, 2005
  \end{center}
  \thispagestyle{empty}
  \cleardoublepage
  \setcounter{page}{1}
  \thispagestyle{empty}

\input{abstract}

\vspace{0.2in}
\sloppypar
\noindent
Keywords:
    Conditional Exponential Family;
    Curved Exponential Family;
    Flat Exponential Family;
    Generalized Linear Model;
    Graphical Model;
 %  Life History Analysis;  %%%% now in title
    Maximum Likelihood;
    Nuisance Variable.

\vspace{0.2in}
\sloppypar
This technical report consists of a draft paper about aster models
supplemented by 5 appendicies on technical subjects.
\begin{itemize}
\item Appendix~\ref{ch:predict} (p.~\pageref{ch:predict})
gives details about ``prediction''
(what the \verb@predict.aster@ function does), although the technical
details are about change-of-parameter formulas and their derivatives.
\item Appendix~\ref{ch:unit} (p.~\pageref{ch:unit})
gives details about the one-parameter exponential families currently
available as conditional distribution of variables given their parent variable
in the \verb@aster@ package.
\item Appendix~\ref{ch:sim} (p.~\pageref{ch:sim})
gives details about simulating a Poisson distribution conditional on being
nonzero (the only non-trivial issue being getting efficient simulation for
unconditional means close to zero).
\item Appendix~\ref{ch:coneflr} (p.~\pageref{ch:coneflr})
gives details of the data analysis that is briefly described
in the draft paper.
\item Appendix~\ref{ch:spaces} (p.~\pageref{ch:spaces})
shows that steepness of conditional exponential families
implies stepness of corresponding unconditional families,
and the full unconditional family gets no new parameter points (that do not
correspond to conditional parameter points).
\end{itemize}

Note added for this git repository.  The text file of the draft paper
that was used to make this technical report seems to have been lost.
(We were not using version control when the paper was written.)
So this document does not reproduce the technical report exactly
(the text here is the text of the published version of the paper).
The PDF of the original technical is found
at \url{https://www.stat.umn.edu/geyer/aster/tr644.pdf}.

  \thispagestyle{empty}
  \cleardoublepage
  \setcounter{page}{1}

\chapter{Draft Paper} \label{ch:aster}

\input{body}

\appendix

\chapter{More Details on Prediction} \label{ch:predict}

\input{predict}

\chapter{Some One-Parameter Exponential Families for Aster Models}
\label{ch:unit}

\input{unit}

\chapter{Simulation of Poisson Conditioned on Being Nonzero} \label{ch:sim}

\input{sim}

\chapter{Example Analysis of Purple Coneflower
    (\emph{Echinacea angustifolia}) Data} \label{ch:coneflr}

\input{coneflr}

\chapter{Canonical Parameter Spaces of Aster Models} \label{ch:spaces}

\input{nats}

\end{document}

