
\documentclass[11pt,twoside,notitlepage]{report}

\usepackage{indentfirst}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{url}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}

\RequirePackage{amsfonts}
\newcommand{\boldbeta}{\boldsymbol{\beta}}
\newcommand{\boldtheta}{\boldsymbol{\theta}}
\newcommand{\boldvarphi}{\boldsymbol{\varphi}}
\newcommand{\boldtau}{\boldsymbol{\tau}}
\newcommand{\boldxi}{\boldsymbol{\xi}}
\newcommand{\boldmu}{\boldsymbol{\mu}}
\newcommand{\boldmuhat}{\boldsymbol{\hat{\mu}}}
\newcommand{\bolda}{\mathbf{a}}
\newcommand{\boldA}{\mathbf{A}}

% original file downloaded from www.stat.umn.edu/geyer/aster/tr658
% had hard-coded path to Sweave.sty that only worked on some computer
% back in 2007
\usepackage{Sweave}

\usepackage{ifthen}
\newboolean{techrept}
\setboolean{techrept}{true}

% remove room for headers, add to textheight

  \addtolength{\textheight}{\headheight}
  \addtolength{\textheight}{\headsep}
  \setlength{\headheight}{0 pt}
  \setlength{\headsep}{0 pt}

% adjust so right and left hand pages have different margins
% not sure why these particular numbers were picked or if they
% still make sense

  % \showthe\evensidemargin
  % \showthe\oddsidemargin
  % \showthe\textwidth

  % \evensidemargin  15 pt
  % \oddsidemargin 61.5 pt
  % \textwidth 392.5 pt

  \evensidemargin 28.75 pt
  \oddsidemargin 75.25 pt
  \textwidth 365 pt

%%%%% NEW go with 1.25 inch margin on all sides

\setlength{\textheight}{\paperheight}
\addtolength{\textheight}{- 2 in}
\setlength{\topmargin}{0.25 pt}
\setlength{\headheight}{0 pt}
\setlength{\headsep}{0 pt}
\addtolength{\textheight}{- \topmargin}
\addtolength{\textheight}{- \topmargin}
\addtolength{\textheight}{- \footskip}

\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{- 2 in}
\setlength{\oddsidemargin}{0.25 in}
\setlength{\evensidemargin}{0.25 in}
\addtolength{\textwidth}{- \oddsidemargin}
\addtolength{\textwidth}{- \evensidemargin}

\begin{document}

  \vspace*{0.9375in}
  \begin{center}
    {\bfseries Supporting Data Analysis for \\
    ``Unifying Life History Analysis for Inference \\
    of Fitness and Population Growth''} \\
    By \\
    Ruth G. Shaw, Charles J. Geyer, Stuart Wagenius, \\
    Helen H. Hangelbroek, and Julie R. Etterson \\
    Technical Report No.~658 \\
    School of Statistics \\
    University of Minnesota \\
%       April 20, 2005 \\
    % original file downloaded from www.stat.umn.edu/geyer/aster/tr658
    % said \today but printed in the PDF
    July 7, 2007
  \end{center}
  \thispagestyle{empty}
  \cleardoublepage
  \setcounter{page}{1}
  \thispagestyle{empty}

\begin{abstract}
This technical report (TR) gives details of the data analyses backing up
a paper \citep{aster2} having the same authors as this TR
and having the title that is quoted in the title of this TR
There are three data sets.
The first data set (Example~1 in the paper, Chapter~\ref{ch:newnew} in
this TR) is new, coming from
Stuart Wagenius's group, and involves the perennial plant
\emph{Echinacea angustifolia} (narrow leaved purple coneflower).
The second data set (Example~2 in the paper, Chapters~\ref{ch:chamae2}
and~\ref{ch:chamae} in this TR) has been previously analyzed
\citep{es,etterson} and involves annual plant \emph{Chamaecrista fasciculata}
(partridge pea).
The third data set (Example~3 in the paper, Chapter~\ref{ch:aphids} in this TR)
has been previously analyzed
\citep{aphids} and involves the insect \emph{Uroleucon rudbeckiae}
(brown ambrosia aphid).

All analyses are done in R \citep{rcore}, all using the \texttt{aster}
contributed package, described by \citet{gws} except for the analyses
in the style of \citet{la}, which use ordinary least squares regression.
All analyses are done using the \texttt{Sweave} function in R, so that
they are completely reproducible by anyone who has R with the \texttt{aster}
package installed and \LaTeX.
\end{abstract}

  \thispagestyle{empty}
  \cleardoublepage
  \setcounter{page}{1}

\include{start}
\include{newnew}
\include{chamae2}
\include{chamae}
\include{aphids}

\begin{thebibliography}{}

\bibitem[Buckheit and Donoho(1995)]{donoho}
Buckheit, J.~B. and Donoho, D.~L. (1995).
\newblock WaveLab and reproducible research.
\newblock \url{http://www-stat.stanford.edu/~donoho/Reports/1995/wavelab.pdf}

\bibitem[Davison and Snell(1991)]{ds}
Davison, A.~C., and Snell, E.~J. (1991).
\newblock Residuals and diagnostics.
\newblock In \emph{Statistical Theory and Modelling: In honour
    of Sir David Cox, FRS\@.}  D.~V. Hinkley, N. Reid, E.~J. Snell (eds.)
\newblock Chapman \& Hall.

\bibitem[Etterson(2004)]{etterson}
Etterson, J.~R. (2004)
\newblock Evolutionary potential of \emph{Chamaecrista fasciculata} in
    relation to climate change.  I. Clinal patterns of selection along
    an environmental gradient in the great plains.
\newblock \emph{Evolution}, \textbf{58}, 1446--1458.

\bibitem[Etterson and Shaw(2001)]{es}
Etterson, J.~R., and Shaw, R.~G. (2001).
\newblock Constraint to adaptive evolution in response to global warming.
\newblock \emph{Science}, \textbf{294}, 151--154.

\bibitem[Gentleman and Temple Lang(2004)]{gent}
Gentleman, R. and Temple Lang, D. (2004).
\newblock Statistical analyses and reproducible research.
\newblock Bioconductor Project Working Papers. Working Paper 2.
\newblock \url{http://www.bepress.com/bioconductor/paper2}

% \bibitem[Geyer(1994)]{g}
% Geyer, C.~J. (1994).
% \newblock On the convergence of Monte Carlo maximum likelihood calculations.
% \newblock \emph{J. Roy.\ Statist.\ Soc.\ Ser.\ B}, \textbf{56}, 261--274.

\bibitem[Geyer, et al.(2007)Geyer, Wagenius and Shaw]{gws}
Geyer, C.~J., Wagenius, S. and Shaw, R.~G. (2007).
\newblock Aster models for life history analysis.
\newblock \emph{Biometrika}, in press.

\bibitem[Lande and Arnold(1983)]{la}
Lande, R. and Arnold, S.~J. (1983).
\newblock The measurement of selection on correlated characters.
\newblock \emph{Evolution}, \textbf{37}, 1210--1226.

\bibitem[Leisch(2002a)]{sweave1}
Leisch, F. (2002a).
\newblock Sweave, part I: Mixing R and LATEX.
\newblock \emph{R News}, \textbf{2}, 28--31.
\newblock \url{http://cran.r-project.org/doc/Rnews}.

\bibitem[Leisch(2002b)]{sweave2}
Leisch, F. (2002b).
\newblock Sweave: Dynamic generation of statistical reports using
    literate data analysis.
\newblock In Wolfgang H\"{a}rdle and Bernd R\"{o}nz, editors,
    \emph{Compstat 2002 --- Proceedings in Computational Statistics},
    pp.~575--580.
\newblock Physica Verlag, Heidelberg.

\bibitem[Lenski and Service(1982)]{aphids}
Lenski, R.~E. and Service, P.~M. (1982).
\newblock The statistical analysis of population growth rates calculated
    from schedules of survivorship and fecunidity.
\newblock \emph{Ecology}, \textbf{63}, 655--662.

\bibitem[McCullagh and Nelder(1989)]{mn}
McCullagh, P., and Nelder, J.~A. (1989).
\newblock \emph{Generalized Linear Models}, 2nd ed.
\newblock Chapman \& Hall.

\bibitem[Mitchell-Olds and Shaw(1987)]{ms}
Mitchell-Olds, T., and Shaw, R.~G. (1987).
\newblock Regression analysis of natural selection: Statistical inference
    and biological interpretation.
\newblock \emph{Evolution}, \textbf{41}, 1149--1161.

\bibitem[R Development Core Team(2006)]{rcore}
\textsc{R Development Core Team} (2006).
\newblock R: A language and environment for statistical computing.
\newblock R Foundation for Statistical Computing, Vienna, Austria.
\newblock \url{http://www.R-project.org}.

\bibitem[Shaw, et al.(submitted) Shaw, Geyer, Wagenius, Hangelbroek, and
    Etterson]{aster2}
Shaw, R.~G., Geyer, C.~J., Wagenius, S., Hangelbroek, H.~H., and
Etterson, J.~R. (submitted).
\newblock Unifying life history analysis for inference of fitness
    and population growth.
\newblock \url{http://www.stat.umn.edu/geyer/aster/}

\bibitem[Stanton and Thiede(2005)]{st}
Stanton, M.~L. and Thiede, D.~A. (2005).
\newblock Statistical convenience vs biological insight: consequences
    of data transformation
    for the analysis of fitness variation in heterogeneous environments
\newblock \emph{New Phytologist}, \textbf{166}, 319--338.

\bibitem[Stein(1981)]{stein}
Stein, C.~M. (1981).
\newblock Estimation of the mean of a multivariate normal distribution.
\newblock {Annals of Statistics}, \textbf{9}, 1135--1151.

% \bibitem[Sung and Geyer(2006)]{sg}
% Sung, Y.~J., and Geyer, C.~J. (2006).
% \newblock Monte Carlo likelihood inference for missing data models.
% \newblock Submitted.
% \newblock \url{http://www.stat.umn.edu/geyer/bernor}.

\end{thebibliography}

\end{document}

